---
title: "Accumulation-curves"
author: "Blanca Arroyo-Correa"
date: "8/8/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)

```

```{r packages, include = FALSE}

library(ggplot2)
library(reshape2)
library(textshape)

library(dplyr)
library(magrittr)

library(vegan)
library(plyr)

```

# Import interaction data 

I used the complete dataset with interactions recorded in La Rocina population for 60 individual plants.

```{r read, echo=FALSE}

rocina <- read.table("../Database/rocina_halimium_09072019.csv", sep=";", header=T)
head(rocina)
str(rocina)
rocina$N_visitations <- as.numeric(rocina$N_visitations)
rocina.fg <- rocina[,c(5,7,8)]
rocina.sp <- rocina[,c(5,6,8)]

```

Next step: creating the interaction matrix, in this case with interactions between individual plants and pollinator functional groups; and including plant traits and fitness.

```{r clean, echo=FALSE}

rocina.m <- dcast(rocina.fg, Plant_ID ~ Pollinator_fg, value.var="N_visitations", fun.aggregate = sum)

#rocina.m[is.na(rocina.m)] <- 0

#plotweb(rocina.m)
#visweb(rocina.m, type="nested")

data <- read.table("../Database/individual_data_halimium_20190714.csv", sep=";", header=T)
data <- merge(rocina.m, data, by="Plant_ID")

ind <- data %>%  
           dplyr::select(c(1, 13:32, 36:39))

ind$Fitness <- ind$Sum_flower_number * ind$Fruit_set * ind$Seed_set

norm <- (ind$Seed_mass - min(ind$Seed_mass)) / (max(ind$Seed_mass) - min(ind$Seed_mass))
ind$Fitness_w <- ind$Sum_flower_number * ind$Fruit_set * ind$Seed_set * norm


# matrix individual plants - pollinator functional groups
rocina.m <- rocina.m %>% column_to_rownames("Plant_ID")
rocina.m$nothing <- NULL
rocina.m$Butterfly <- NULL

## matrix individual plants - pollinator species 
rocina.m.sp <- dcast(rocina.sp, Plant_ID ~ Pollinator_sp, value.var="N_visitations", fun.aggregate = sum)
rocina.m.sp <- rocina.m.sp %>% column_to_rownames("Plant_ID")
rocina.m.sp$nothing <- NULL
rocina.m.sp$Butterfly <- NULL



```


# Accumulation curves

In order to fit the number of censuses to the local abundance of flower visitors in the population I constructed accumulation curves using the ‘vegan’ package. I first built the accumulation curve for the pollinator functional groups assemblage, where x axis represent the number of plants recorded and y de pollinator f.g. detected. Second, I performed the same analysis for each of the individual plants. 

```{r accum}

# accumulation curve with pollinator composition (not interactions) (x= plants recorded, y= pollinator functional groups detected)

hal.int <- specaccum(rocina.m, method = "random")

par(mar=c(5,4,4,4))
plot(hal.int, ci.type = "poly", col = "#4E9C7B", lwd = 2, ci.lty = 0, ylim = c(0, 13), 
    ci.col = alpha("#4E9C7B", alpha = .4), xlab = "Number of individual plants surveyed", ylab = "Number of pollinator functional groups", cex.axis=1.5, cex.lab=1.5)
boxplot(hal.int, col = "orange", add = TRUE, pch = "+")

est.richness<-fossil::chao1(rocina.m,taxa.row = FALSE)

(complet<-ncol(rocina.m)*100/est.richness)

specpool(rocina.m)
fossil::spp.est(rocina.m)

# accumulation curve with pollinator composition (not interactions) (x= plants recorded, y= pollinator species detected)

hal.int.sp <- specaccum(rocina.m.sp, method = "random")

par(mar=c(5,4,4,4))
plot(hal.int.sp, ci.type = "poly", col = "#4E9C7B", lwd = 2, ci.lty = 0, ylim = c(0, 25), 
    ci.col = alpha("#4E9C7B", alpha = .4), xlab = "Number of individual plants surveyed", ylab = "Number of pollinator species", cex.axis=1.5, cex.lab=1.5)
boxplot(hal.int.sp, col = "orange", add = TRUE, pch = "+")

est.richness<-fossil::chao1(rocina.m.sp,taxa.row = FALSE)

(complet<-ncol(rocina.m.sp)*100/est.richness)

specpool(rocina.m.sp)
fossil::spp.est(t(rocina.m.sp))


# Individual plants (only 161 for now as an example) --> I'm not sure if this is OK

rocina.ind <- read.table("../Database/rocina_halimium_09072019.csv", sep=";", header=T)
rocina.ind <- dcast(rocina.ind, ID + Plant_ID ~ Pollinator_fg, value.var="N_visitations")
rocina.ind <- rocina.ind %>% column_to_rownames("ID")
rocina.ind$nothing <- NULL

rocina.ind[is.na(rocina.ind)] <- 0

# subset each individual into its own df (example with one individual plant)
rocina.ind %>% filter(Plant_ID == "161") -> p161
droplevels(p161) -> p161

p161 <- p161[, colSums(p161 != 0) > 0] # remove columns with 0 values

plot(specaccum(p161[, 2:5], method = "random"), ci.type = "poly", col = "blue", 
     lwd = 2, ci.lty = 0, ylim = c(0, 7),  ci.col = "lightblue", xlab = "Number of interactions recorded", 
     ylab = "Number of unique interactions")
boxplot(specaccum(p161[, 2:5], method = "random"), col = "yellow2", add = TRUE, 
        pch = "+")

est.richness<-fossil::chao2(p161[, 2:5],taxa.row = FALSE)
(complet<-ncol(p161[, 2:5])*100/est.richness)


#### in a loop for all the individual plants

for(i in seq_along(levels(rocina.ind$Plant_ID))){
  if(i != length(levels(rocina.ind$Plant_ID))){
  temp <-  rocina.ind %>% filter(Plant_ID %in% c(levels(Plant_ID)[i], levels(Plant_ID)[i])) %>% print() %>%
      data.frame()
  temp <- droplevels(temp)
  temp2 <- temp[, colSums(temp != 0) > 0]
  temp2$Plant_ID <- NULL
  assign(paste("p", levels(rocina.ind$Plant_ID)[i], sep = "_"), temp2)
  }
}


l.df <- lapply(ls(pattern="p_"), function(x) get(x))
ls_names <- ls(pattern = "p_.*")
names(l.df) <- ls_names

for(i in 1:length(l.df)){
  x <- str(l.df[[i]])
  print(x)
  }

for(i in 1:length(l.df)){
  est.richness<-fossil::chao2(l.df[[i]],taxa.row = FALSE)
  complet<-ncol(l.df[[i]])*100/est.richness
  print(complet)
  }


```
